
### DOWNLOAD EXTENDED DATA

After we have imported the libraries we can download the extended data that we gathered from our spotify download. Since all the files I need are .json I use the pattern *json to load them all in at once instead of individually loading them in.
```{r}
library(tidyverse)
library(jsonlite)
library(lubridate)
library(plotly)
library(flexdashboard)
library(spotifyr)
library(purrr)
library(knitr)
```

```{r, echo=TRUE}
path <- "./my_extended_data/"
files <- dir(path, pattern = "*.json")

#data <- files %>%
 #      map_df(~fromJSON(file.path(path, .), flatten = TRUE))
data <- read_csv("./my_extended_data/my_data.csv")
```

>To get a better understanding of what all the data fields are make sure to look through the [Start Here](#page-2) page

We can also use glimpse to take a quick look at the data to see what type each column is 
```{r, echo=TRUE}
glimpse(data)
```



After looking through the columns I want to save the ones that I will be using for the analysis.
```{r, echo=TRUE}
spotify_data <- data %>% 
  select(ts,ms_played,master_metadata_track_name,master_metadata_album_artist_name,master_metadata_album_album_name,reason_start,reason_end)
```

### Data Wrangling Date Time & Rename Column

Throughout the semester I learned how important tidy data is. Here we will impliment two of the changes to make this dataframe easier to manage.

* Converting a column from char to POSIXct. This will allow for easier data analysis since the column will be of type date. We can use the `lubridate` to gain more information.

* Renaming columns for easier analysis 


```{r, echo=TRUE}
spotify_data <- spotify_data %>% 
  as_tibble() %>%
  separate(col = "ts", 
           into = c("date","time"),
           sep = "T") %>% 
  separate(col = "time",
           into = "time",
           sep = "Z")

```


```{r, echo=TRUE}
datetime<- as.POSIXct(paste(spotify_data$date, spotify_data$time), format = "%Y-%m-%d %H:%M:%S")
```

```{r, echo=TRUE}
spotify_data$datetime <- datetime
```

Here we can check that the change was implimented 
```{r, echo=TRUE}
spotify_data %>% 
  select(datetime)
```


Finally, we just need to rename the columns.
```{r, echo=TRUE}
spotify_data <- spotify_data %>% 
  rename('track'='master_metadata_track_name',
         'artist'='master_metadata_album_artist_name',
         'album' ='master_metadata_album_album_name'
         )
```

>Now that we have cleaned up the data a bit we can move into data exploration and analysis! 

### Total Songs Played 

I've used this spotify account since 2018. I know I really enjoy listening to music all the time. The first insight I want to gain is on which day of the week do I listen to the most music.
```{r, echo=TRUE}
library(lubridate)
viz <- spotify_data %>%
  drop_na(datetime) %>% 
  mutate(weekday = wday(datetime, label =TRUE)) %>% 
  group_by(weekday) %>% 
  count(weekday, na.rm=T) %>% 
  ggplot(aes(weekday, n)) +
  geom_point() +
  labs(
    x = "day of the week",
    y = "number of songs played from 2018-2023",
  )
```

It looks like my Monday's typically consist of a cup of coffee and some music to start my week off right! 

### total songs played during week plot  {data-height=600}

```{r, fig.align='left', echo = FALSE}
viz
```



### Top 5 favorite songs through the years

I also wanted to learn about what my favorite songs were throughout the years. Since I tend to listen to a lot of music I only want to display my top 5 results.

This shows my top 5 songs throughout the years
```{r, echo=TRUE}
 top_five_songs<-spotify_data %>% 
   mutate(year= year(datetime)) %>% 
   na.omit() %>% 
   group_by(year,track) %>%
   summarise(n=n()) %>% 
   top_n(n = 5,wt=n) 
top_five_songs

```

```{r, echo=TRUE}
top_songs_bar <- ggplot(top_five_songs, aes(fill=track, x=year,y=n))+
  geom_bar(position="stack", stat="identity",show.legend = FALSE)+
  scale_x_continuous(breaks = c(2018:2023))
```

**Question for REI How do i make the chart fit the year is not showing** 

### Visualization {data-height=750, data-width=500}

```{r,fig.align='left',echo=TRUE}
ggplotly(top_songs_bar+ labs(title = "Top 5 Songs Throughout The Years")) %>% 
  layout()

```



### Top songs no skip

The analysis for my top 5 tracks tracked every time the song was played. However, just because a song plays doesn't mean I listened to it all the way through. I will recreate what I did for the top tracks but this time i will only count the songs that played all the way through.


```{r, echo=TRUE}
skipped <-spotify_data %>% 
 #select(9:11,17,23) %>% 
 # na.omit() %>% 
  filter(reason_end == "trackdone") %>%
  mutate(year= year(datetime)) %>% 
  group_by(year,track) %>%
   summarise(n=n()) %>% 
   top_n(n = 5,wt=n) 
```

```{r}
skipped
```


```{r, echo=TRUE}
done <- ggplot(skipped, aes(fill=track, x=year,y=n))+
  geom_bar(position="stack", stat="identity",show.legend = FALSE)+
  scale_x_continuous(breaks = c(2018:2023))
```

### Visualization {data-height=750}

```{r, echo=TRUE}
done_plotly <- ggplotly(done+ labs(title = "Top 5 Songs Throughout The Years - Track Done")) %>% 
  layout(showlegend = FALSE)

done_plotly
```


### How often did I listen to my top song 

**Once i filtered by the songs that finished playing without me skipping you can see that my top 5 tracks change throughout the year**

my top song in 2019 was Bow Down from Iprevail played a total of 116 times. Which day of the week did i listen to the song the most

```{r, echo=TRUE}
amount_played_2019 <- spotify_data %>% 
  drop_na(datetime) %>% 
  mutate(weekday = wday(datetime, label =TRUE),
         year = year(datetime)
         )%>% 
  select(weekday,year,track,datetime) %>% 
  na.omit() %>% 
  group_by(track,weekday,year) %>% 
  count() %>% 
  filter(track == "Bow Down", year == 2019)
```

```{r}
amount_played_2019
```

This table shows that in 2019 on in total on Mondays I listened to Bow Down the most. Which makes sense becasue I typically went to the gym on monday mornings that yearr. 

### How many tracks do I skip a year?

After seeing how my top 5 changed when i filtered out by songs that finished. I was curious to see how many songs I skip every year.

```{r, echo=TRUE}
skipped_yearly <- spotify_data %>% 
   mutate(year = year(datetime))%>% 
  select(track,reason_end,year) %>% 
  na.omit() %>% 
  group_by(year,reason_end) %>% 
  count()
```

```{r}
skipped_yearly
```

```{r}
skipped_yearly_plot <- ggplot(data= skipped_yearly, aes(x= year, y=n, fill=reason_end))+
  geom_point()
```

### Visualization {data-height=750}

This is the rendered graphic showing how the songs I listened to ended throughout the years.

```{r}
ggplotly(skipped_yearly_plot)
```


### Total listening hours


I know I spend a lot of time listening to music so I wanted to visualize this. 


```{r, echo=TRUE}
year_plot <-spotify_data %>%
  mutate(date = date(datetime)) %>% 
  group_by(date = floor_date(date, "week")) %>% 
  summarise(total_ms_played = sum(ms_played)) %>% 
  na.omit()%>% 
  mutate(second = (total_ms_played/ 1000),
         minute = (second/60),
         hours = (minute/60))  %>% 
  ggplot(aes(x = date, y = hours)) +
  geom_col(aes(fill = hours)) +
  scale_x_date(date_labels = "%b %Y", date_breaks= "12 month") +
  scale_fill_gradient(low = "grey", high = "gold") +
  labs(title = "Playback hours per week") +
  xlab("") +
  theme_light()
```

### Visualization {data-height=600} 
 The bar graph below shows how many hours a **week** I spent listening to music.

```{r}
ggplotly(year_plot+ labs(title = "Hours spent listening to music weekly")) 

```

I was curious to see what I was doing at that time... I had a couple of road trips so it makes sense that my listening hours were at an all time high! 


### Breakdown of the amount of songs I listened to that week

Because the highest week I had of listening to music I was curious to see how many songs I listened to each day 

```{r}
week_of_oct <-spotify_data %>%
  mutate(date = date(datetime)) %>% 
  select(ms_played,track,date) %>% 
  filter(date %in% ymd(c("2020-10-25","2020-10-26","2020-10-27","2020-10-28","2020-10-29","2020-10-30","2020-10-31"))) %>% 
  group_by(track,date) %>% 
  mutate(amount_played = n()) %>% 
  group_by(date) %>% 
  summarise(
    total_play = sum(amount_played)
  )


```

```{r}
week_of_oct_plot <- ggplot(data=week_of_oct, aes(x=date, y=total_play))+
  geom_bar(position="stack", stat="identity",show.legend = FALSE)+
  scale_x_date(date_labels = "%b %D", date_breaks= "1 day") +
    theme(axis.text = element_text(angle = 90, vjust = 0.5, hjust = 1))


```

```{r}
ggplotly(week_of_oct_plot) %>% 
  layout(showlegend = FALSE)

```

Using spotifyWebAPI it's harder to get the same insights that i got from my static analysis 

### Listening History from the start of my spotify account 

After you follow the steps from on how to get your client id ([Start Here](#page-1)) we can get started on our dynamic analysis! 

This portion uses a lot of the `spotifyr` package so make sure you ran the first import libraries 

first we want to reference our credentials. We make sure this is not anywhere visible so i saved it in a new file. 


```{r}
source("keys/spotifyAPI.R")
```


this grants our authorization so we can use the web api.
```{r}
# auth<- spotifyr::get_spotify_authorization_code(Sys.getenv("SPOTIFY_CLIENT_ID"),
#                                         Sys.getenv("SPOTIFY_CLIENT_SECRET"),
#                                         scopes)
```

After running this command you'll get a new window from spotify asking for your credentials 

### What Are My Favorite Tracks Within the Past 4 Weeks

Using the spotifyAPI you can get different analysis 

```{r, echo=TRUE}
top_trackATM <- get_my_top_artists_or_tracks(type = 'tracks', 
                             time_range = 'short_term', 
                             limit = 5) %>% 
    mutate(
        artist.name = map_chr(artists, function(x) x$name[1])
        ) %>% 
  select(artist.name,name,popularity)
```

Using the spotify API we can get more information on 

### Vizualization

Lets see how popular my top most recent favorite tracks are
```{r}
ggplot(data = top_trackATM, aes(x = name, y = popularity)) + 
  coord_flip()+
  geom_col(aes(fill = artist.name))
```


### Favorite track within the past 6 months 

```{r, echo=TRUE}
top_track_6month <- get_my_top_artists_or_tracks(type = 'tracks', 
                             time_range = 'medium_term', 
                             limit = 5,
                             ) %>% 
    mutate(
        artist.name = map_chr(artists, function(x) x$name[1])
        ) %>% 
  select(artist.name,name,popularity)
```

I wasn't expecting _just pretend_ to have the highest popularity score since it's more of a rock song.


### Vizualization

```{r}
ggplot(data = top_track_6month, aes(x = name, y = popularity)) +
  coord_flip()+
  geom_col(aes(fill = artist.name)) 
```



### What Type of Features Can I Get From My Playlist


```{r}
playlist <- get_my_playlists()
```


***EXPLAIN WHAT ENERGY & DANCE IS***
B2B is a playlist with artist from recent/upcomoming concerts 

```{r, echo=TRUE}
b2b_audio <- get_playlist_audio_features(playlist_uris = "69LURFh6OXwpXQWSQc45e7")
```

```{r, echo=TRUE}
b2b_audio_plot <- ggplot(b2b_audio, aes(fill = track.name, y= danceability, x = tempo)) +
  geom_point() 

```

```{r}
b2b_audio %>% 
      mutate(
        artist.name = map_chr(track.artists, function(x) x$name[1])
        ) %>% 
  select(track.name,artist.name,danceability,energy) %>% 
  arrange(desc(danceability))
```


### Visualization {data-height=800} 
```{r, echo=TRUE}
ggplotly(b2b_audio_plot+ labs(title = "Danceability of B2B Playlist")) 
  
```
One big difference from what I did in my extended history analysis is that with the API i can get more information about tracks

## What More Can I Learn About My 2019 Favorite Songs 
```{r}
playlist
```

```{r}
playlist_2019 <- get_playlist_audio_features(playlist_uris = "37i9dQZF1EtklzPPpNlUF9") 
  

```

```{r}
playlist_2019 %>% 
  group_by(track.artists)
```

```{r}
 playlist_2019 <- playlist_2019%>% 
      mutate(
        artist.name = map_chr(track.artists, function(x) x$name[1])
        ) %>% 
  select(artist.name,track.name, key_name, key_mode,danceability,energy,loudness,liveness,valence) 
  
  
```

```{r}
plot_2019 <- ggplot(data=playlist_2019, aes(x=loudness, y=energy, fill = track.name))+
  geom_point(show.legend = F)
```

```{r}
ggplotly(plot_2019)
```

